---
title: "Build Predictive Model for Loan Prediction Data Set"
author: "Jake Bersabe"
date: "08/12/2021"
output: 
  html_document:
    css: "style.css"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F, fig.align = "center")
```

## Import the data set and libraries

```{r}
library(dplyr)

# library(rsample)
# library(parsnip)
# library(yardstick)
# library(recipes)
# library(workflows)

library(tidymodels)
library(themis)

data <- readr::read_csv("train.csv")
glimpse(data)
```

## Create `recipe()`

### Prepare Data for `recipe()`

```{r}
# Function to add monthly payment column
monthly_payment <- function(loan_amount, loan_amount_term){
  r <- 0.005
  monthly_payment <- (loan_amount/(((1+r)^loan_amount_term)-1))*(((1+r)^loan_amount_term)*r)
  return(monthly_payment)
} 


data_4_rec <- data %>%
  mutate(Gender = factor(Gender),
         Married = factor(Married),
         Education = factor(Education),
         Self_Employed = factor(Self_Employed),
         Property_Area = factor(Property_Area)) %>% 
  mice::mice(m=2, maxit = 2, method = 'cart', seed = 500) %>% 
  mice::complete(1) %>% 
  mutate(LoanAmount = LoanAmount*1000) %>% 
  select(-Loan_ID)


data_4_rec %>% glimpse()
colSums(is.na(data_4_rec))
```



### Split Data and Folds

```{r}
set.seed(123)
data_split <- initial_split(data_4_rec, strata = Loan_Status)
data_tr <- training(data_split)
data_te <- testing(data_split)

data_tr %>% glimpse()

set.seed(345)
folds <- vfold_cv(data_tr)
folds
```

### The Recipe, Finally

```{r}
data_rec <- recipe(Loan_Status ~ ., data = data_tr) %>% 
  step_string2factor(Loan_Status) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_smote(Loan_Status, seed = 123) %>% 
  step_zv(all_predictors()) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())

```


Let's take a glimpse at the pre-processed data.
```{r}
data_rec %>% prep() %>% bake(new_data = NULL) %>% glimpse()
```



Notice the class imbalance in the target variable:
```{r}
data_tr %>% 
  group_by(Loan_Status) %>% 
  count()
```

The class imbalance is resolved by `step_smote()`
```{r}
data_rec %>% prep() %>% bake(new_data = NULL) %>% count(Loan_Status)
```

## Build Model

### Create Model Specification

```{r}
glm_spec <- logistic_reg() %>% 
  set_engine("glm")
```


### Create **Workflow**

```{r}
data_wf <- workflow() %>% 
  add_recipe(data_rec) %>% 
  add_model(glm_spec)

data_wf
```

## Fit Model

### Fit to Folds

```{r}
glm_rs <- data_wf %>% 
  fit_resamples(
    resamples = folds,
    metrics = metric_set(roc_auc, accuracy, sensitivity, specificity),
    control = control_resamples(save_pred = T)
  )

glm_rs
```

### Collect Metrics

```{r}
collect_metrics(glm_rs)
```

```{r}
glm_rs %>% 
  conf_mat_resampled()
```

### Plot ROC Curve

```{r}
glm_rs %>% 
  collect_predictions() %>% 
  group_by(id) %>% 
  roc_curve(Loan_Status, .pred_N) %>% 
  autoplot()
```

## Final Fit

```{r}
final_fit <- data_wf %>% 
  last_fit(data_split,
           metrics = metric_set(roc_auc, accuracy, sensitivity, specificity, kap),
    control = control_resamples(save_pred = T))

final_fit %>% head(10)
```

```{r}
collect_metrics(final_fit)
```

```{r}
final_fit %>% 
  collect_predictions() %>% 
  group_by(id) %>% 
  roc_curve(Loan_Status, .pred_N) %>% 
  autoplot()
```

```{r}
collect_predictions(final_fit) %>% 
  conf_mat(Loan_Status, .pred_class)
```

```{r}
final_fit %>% 
  pull(.workflow) %>% 
  pluck(1) %>% 
  tidy(exponentiate = T) %>% 
  arrange(desc(estimate)) %>% 
  filter(term != "(Intercept)")
```


```{r}
final_fit %>% 
  pull(.workflow) %>% 
  pluck(1) %>% 
  tidy() %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(estimate, forcats::fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, size = 1.2) +
  geom_errorbar(aes(xmin = estimate - std.error,
                    xmax = estimate + std.error),
                width = 0.2, alpha = 0.7) +
  geom_point()+
  labs(x = "Estimate",
       y = "")
```

The variables to the right of the vertical line increase the probability of Y in Loan_Status and those to the left does the opposite.
