---
title: "Preprocessing Loan Eligibility Prediction"
author: "Jake Bersabe"
date: "14/11/2021"
output: 
  html_document:
    css: "style.css"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F, fig.align = "center")
```

## Introduction

### Import the data set  

```{r message=FALSE}
library(dplyr)
library(printr)
data <- readr::read_csv("train.csv")
glimpse(data)
```

```{r message=FALSE}
library(skimr)
skim(data)
```  


### Split data set  
```{r}
library(rsample)

data_xplit <- initial_split(data, strata = Loan_Status)

data_train <- training(data_xplit)
data_test <- testing(data_xplit)

```



## Preprocessing 1: Step by step
### 1. Imputation, deselect LoanID   
```{r message=FALSE}
library(mice)

data_train1 <- data_train %>% 
  mutate(Gender = factor(Gender),
         Married = factor(Married),
         Education = factor(Education),
         Self_Employed = factor(Self_Employed),
         Property_Area = factor(Property_Area)) %>% 
  select(-Loan_ID) %>% 
  mice(m=2, maxit = 2, method = 'cart', seed = 500) %>% 
  complete(1)

colSums(is.na(data_train1))
```

### 2. Feature engineering:  Add CombinedIncome column, Deselect applicant and coapplicant income, add MonthlyPayment column, PaymentIncomeRatio column 

First, create a function that will compute monthly payment.  
```{r}
monthly_payment <- function(loan_amount, loan_amount_term){
  r <- 0.005
  monthly_payment <- (loan_amount/(((1+r)^loan_amount_term)-1))*(((1+r)^loan_amount_term)*r)
  return(monthly_payment)
}
```



```{r}
data_train2 <- data_train1 %>% 
  mutate(LoanAmount = LoanAmount*1000,
         CombinedIncome = ApplicantIncome + CoapplicantIncome) %>% 
  select(-ApplicantIncome, -CoapplicantIncome) %>% 
  mutate(MonthlyPayment = monthly_payment(LoanAmount, Loan_Amount_Term),
         PaymentIncomeRatio = MonthlyPayment/CombinedIncome)
glimpse(data_train2)
```




### 3. Dummy Coding  
```{r}
data_train3 <- data_train2 %>% 
  mutate(Gender = if_else(Gender == "Male", 1, 0),
         Married = if_else(Married == "Yes", 1, 0),
         Education = if_else(Education == "Graduate", 1, 0),
         Self_Employed = if_else(Self_Employed == "Yes", 1, 0),
         Credit_History = as.numeric(as.character(Credit_History)),
         Property_Area_Semi = if_else(Property_Area == "Semiurban", 1, 0),
         Property_Area_Rural = if_else(Property_Area == "Rural", 1, 0),
         Property_Area_Urban = if_else(Property_Area == "Urban", 1, 0),
         Loan_Status = if_else(Loan_Status == "Y", 1, 0)) %>% 
  select(-Property_Area)
data_train3 %>% glimpse()
```


### 4. Normalization  
```{r}
data_train4 <- scale(data_train3) %>% as_tibble()
glimpse(data_train4)
```

```{r}
library(corrplot)
data_train4 %>% 
  select(-LoanAmount, -Loan_Amount_Term, -CombinedIncome, -MonthlyPayment) %>% 
  cor() %>% 
  corrplot(type = "lower")
```


## Preprocessing 2: Pipeline  
```{r}
# Library for imputation
library(mice) 

# Function to add monthly payment column
monthly_payment <- function(loan_amount, loan_amount_term){
  r <- 0.005
  monthly_payment <- (loan_amount/(((1+r)^loan_amount_term)-1))*(((1+r)^loan_amount_term)*r)
  return(monthly_payment)
} 

# Pre-processing pipeline 

data_train_pipelined <- data_train %>%
  mutate(Gender = factor(Gender),
         Married = factor(Married),
         Education = factor(Education),
         Self_Employed = factor(Self_Employed),
         Property_Area = factor(Property_Area)) %>% 
  select(-Loan_ID) %>% 
  mice(m=2, maxit = 2, method = 'cart', seed = 500) %>% 
  complete(1) %>% 
  mutate(LoanAmount = LoanAmount*1000,
         CombinedIncome = ApplicantIncome + CoapplicantIncome) %>% 
  select(-ApplicantIncome, -CoapplicantIncome) %>% 
  mutate(MonthlyPayment = monthly_payment(LoanAmount, Loan_Amount_Term),
         PaymentIncomeRatio = MonthlyPayment/CombinedIncome) %>% 
  mutate(Gender = if_else(Gender == "Male", 1, 0),
         Married = if_else(Married == "Yes", 1, 0),
         Education = if_else(Education == "Graduate", 1, 0),
         Self_Employed = if_else(Self_Employed == "Yes", 1, 0),
         Credit_History = as.numeric(as.character(Credit_History)),
         Property_Area_Semi = if_else(Property_Area == "Semiurban", 1, 0),
         Property_Area_Rural = if_else(Property_Area == "Rural", 1, 0),
         Property_Area_Urban = if_else(Property_Area == "Urban", 1, 0),
         Loan_Status = if_else(Loan_Status == "Y", 1, 0)) %>% 
  select(-Property_Area) %>% 
  scale() %>% 
  as_tibble()
  
glimpse(data_train_pipelined)
```


## Compare for Difference
```{r}
waldo::compare(data_train_pipelined, data_train4)
```

## Create `recipe()`

### Prepare Data for `recipe()`
```{r}
data_4_rec <- data %>%
  mutate(Gender = factor(Gender),
         Married = factor(Married),
         Education = factor(Education),
         Self_Employed = factor(Self_Employed),
         Property_Area = factor(Property_Area)) %>% 
  mice(m=2, maxit = 2, method = 'cart', seed = 500) %>% 
  complete(1) %>% 
  mutate(LoanAmount = LoanAmount*1000,
         CombinedIncome = ApplicantIncome + CoapplicantIncome) %>% 
  mutate(MonthlyPayment = monthly_payment(LoanAmount, Loan_Amount_Term),
         PaymentIncomeRatio = MonthlyPayment/CombinedIncome) %>% 
  select(-ApplicantIncome, -CoapplicantIncome, -LoanAmount, -Loan_Amount_Term, -Loan_ID)


data_4_rec %>% glimpse()
```

### Split Data
```{r}
set.seed(123)
data_split <- initial_split(data_4_rec, strata = Loan_Status)
data_tr <- training(data_split)
data_te <- testing(data_split)
```


### The Recipe, Finally
```{r}
data_tr %>% glimpse()

library(recipes)
data_rec <- recipe(Loan_Status ~ ., data = data_tr) %>% 
  step_string2factor(Loan_Status) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())
  
```

### `prep()`
```{r}
data_prep <- data_rec %>% prep()
bake(data_prep, data_te) %>% glimpse()
```





















